\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-18 doc/paper.Rmd
alternative different shifts

Same-shift pairs of length $w_v$ are defined in \citet{hadler} as all 
pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i}^{(2)}$ are defined (see \autoref{sketch-same-diff}).

@@

\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). }
\end{figure}

@@

The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e. $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$. For uniformly distributed optimal locations, we expect to see these situation with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$. For  an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.

The number of possible different-shift pairs also depends on the location of  the  optimal locations $t_1^o$ and $t_2^o$. Whenever the optimal locations are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$. For uniformly distributed locations $t_i^{(o)}$, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.

\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
\hh{figure: number of diff shift pairs and location}

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
\end{figure}

>>>>

Same-shift pairs of length $w_v$ are defined in \citet{hadler} as all 
pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i-1}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \autoref{sketch-same-diff}).

@@

\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
\end{figure}

@@

The number of same-shift pairs will be zero, if the optimal locations $t_1^{o}$ and $t_2^{o}$ are so far apart, that no segments of size $w_v$ are left on the same sides of the optimal locations, i.e. $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or  $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$. While we can assume that once (sub-)class characteristics are removed, optimal locations $t_i^{o}$ are uniformly distributed across the length of the profile, we cannot assume that $t_1^o$ and $t_2^o$ are independent of each other. In particular, for same-source profiles, we would expect a strong dependency between these locations. However, for different source matches, an argument of independence in the locations could be made. In that case, we expect a test to fail with a probability of $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$. For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.

The number of possible different-shift pairs also depends on the location of  the  optimal locations $t_1^o$ and $t_2^o$. Whenever the optimal locations are close to the boundaries, the number of possible pairings decreases and reaches zero, if $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$. Assuming a correlation between optimal locations $t_1^o$ and $t_2^o$ of close to one for same-source profiles, this results in an expected rate of failure of $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels. Assuming independence in the optimal locations for different source profiles the expected probability for a failed test is, again, $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$.

While failures due to missing correlations from same-shift pairs are unavoidable by definition of the Chumbley score, failures due to missing correlations from different-shift pairs can be prevented by using a different strategy in assigning pairs.

Using the same notation as in \autoref{eq.start}, we define same-shift pairs identical to \citet{hadler} as pairs $(s_i^{(1)}, s_i^{(2)})$ for all $i$ where the boundary conditions of both sequences are met simultaneously. Let us assume that this results in $I$ pairs. Define $s_(j)^{(k)}$ to be the $j$th starting location in sequence $k = 1, 2$, i.e.\ $s_(1)^{(k)} < s_(2)^{(k)} < ... < s_(I)^{(k)}$.

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}
i.e.\ for an odd number of same-shift correlations, we skip the middle pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the same or at most one less than the number of same-shift pairings in all tests.

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
\end{figure}

### 2018-04-18 doc/paper.tex
alternative different shifts

Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
pairs \((s_i^{(1)}, s_i^{(2)})\) for all integer values \(i\) for which
both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
different-shift pairs are defined as \((s_i^{(1)}, s_{-i}^{(2)})\) for
all \(i\) where both \(s_i^{(1)}\) and \(s_{-i}^{(2)}\) are defined (see
\autoref{sketch-same-diff}).

@@

@@

Looking closer, we see that the problem lies in the way the
different-shift pairs are picked: the number of possible different-shift
pairs depends on the location of the optimal locations \(t_1^o\) and
\(t_2^o\). Whenever are close to the boundaries, the number of possible
pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
\(t_i^{(o)} > T_i-w_o- w_v\). For uniformly distributed locations
\(t_i^{(o)}\), this results in a probability of failure of
\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels.

\hh{diagnose problem: use of optimal location in getting the pairs for different-shift correlations}
\hh{figure: number of diff shift pairs and location}

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. }
\end{figure}

>>>>

Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
pairs \((s_i^{(1)}, s_i^{(2)})\) for all integer values \(i\) for which
both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
(see \autoref{sketch-same-diff}).

@@

\subsection{A problem with failures and
modification}\label{a-problem-with-failures-and-modification}

Looking closer at \autoref{eq.start}, we see that by definition, some
number of tests will fail to produce a result, either because the number
of eligible same-shift pairs is zero, or the number of different-shift
opairs is zero.

The number of same-shift pairs will be zero, if the optimal locations
\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
\(w_v\) are left on the same sides of the optimal locations, i.e.
\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\). While we can assume
that once (sub-)class characteristics are removed, optimal locations
\(t_i^{o}\) are uniformly distributed across the length of the profile,
we cannot assume that \(t_1^o\) and \(t_2^o\) are independent of each
other. In particular, for same-source profiles, we would expect a strong
dependency between these locations. However, for different source
matches, an argument of independence in the locations could be made. In
that case, we expect a test to fail with a probability of
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
probability is about 0.0015.

The number of possible different-shift pairs also depends on the
location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
optimal locations are close to the boundaries, the number of possible
pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
profiles, this results in an expected rate of failure of
\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
independence in the optimal locations for different source profiles the
expected probability for a failed test is, again,
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).

While failures due to missing correlations from same-shift pairs are
unavoidable by definition of the Chumbley score, failures due to missing
correlations from different-shift pairs can be prevented by using a
different strategy in assigning pairs.

Using the same notation as in \autoref{eq.start}, we define same-shift
pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
for all \(i\) where the boundary conditions of both sequences are met
simultaneously. Let us assume that this results in \(I\) pairs. Define
\(s_(j)^{(k)}\) to be the \(j\)th starting location in sequence
\(k = 1, 2\), i.e.~\(s_(1)^{(k)} < s_(2)^{(k)} < ... < s_(I)^{(k)}\).

We then define the pairs for different-shifts as

\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation}

i.e.~for an odd number of same-shift correlations, we skip the middle
pair for the different-shift correlations (see \autoref{sketch-diff-2}).
This pairing ensures that the number of different-shift pairings is the
same or at most one less than the number of same-shift pairings in all
tests.

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
\end{figure}

@@
